---
title: "Analysis on Palmer Penguins dataset"
author: "Chiejine Peace"
date: "2023-03-04"
output: pdf_document

---


```{r, echo = FALSE}
knitr::opts_chunk$set (error = TRUE)
```

# Data analysis on the Palmerpenguins data set

## Setting up my environment

Notes: setting up my environment by loading the *tidyverse* and *palmerpenguins* packages

```{r}
install.packages("tidyverse")
library(tidyverse)
install.packages("palmerpenguins")
library(palmerpenguins)
```

## View the data set

Notes: View the data set and note column names

```{r}
View(penguins)
```

## Transform data set

Notes: Transform columns *body_mass_g* and *flipper_length_mm* to new columns measured in kilograms and meters respectively. Name new columns *body_mass_kg* and *flipper_length_m*

```{r}
penguins_viz <- penguins %>% 
  select(species, sex, body_mass_g, flipper_length_mm) %>% 
  mutate(body_mass_kg = body_mass_g / 1000, flipper_length_m = flipper_length_mm / 1000)
head(penguins_viz)
```

## Visualization 1

Scatterplot (*with trendline*) chart to show the relationship between body mass(kg) and flipper length (m) '

```{r}
ggplot(penguins_viz) + 
  geom_smooth( mapping = aes(x =flipper_length_m, y = body_mass_kg)) +
  geom_point( mapping = aes(x =flipper_length_m, y = body_mass_kg, shape = species))
```

## Visualization 2

Notes: Line chart visualization to show relationship between flipper length and body_mass for distinct species of penguins

```{r}

ggplot(penguins_viz) +
  geom_smooth(mapping = aes(x =flipper_length_m, y = body_mass_kg, linetype = species)) + 
  facet_wrap(~species)

```

## Visualization 3

Notes: Facet wrap to show relationship between flipper length and body mass for distinct species and sex.

```{r}
ggplot(penguins_viz) +
  geom_point(mapping = aes(x =flipper_length_m, y = body_mass_kg, color = species)) + 
  facet_grid(~sex~species)

```

## Conclusion

There is a relationship between body mass and flipper length for all species of penguins withe Gentoo specie being the largest and having longer flipper lengths than the other species.
